---
title: "Part2: Basic Inferential Data Analysis Instructions"
output: html_document
---
## Intro
To analyze the ToothGrowth data in the R datasets package.
1. Load the ToothGrowth data and perform some basic exploratory data analyses
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
4. State your conclusions and the assumptions needed for your conclusions.
```{r setup, include=FALSE, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: load the ToothGrowth data

```{r ToothGrowth }
library(ggplot2)
data(ToothGrowth)
```

## Step 2: Basic Summary of ToothGrowth 
```{r summary}
summary(ToothGrowth)
str(ToothGrowth)
head(ToothGrowth)
dose05 <- subset(ToothGrowth,dose==0.5)
dose10 <- subset(ToothGrowth, dose==1.0)
dose20 <- subset(ToothGrowth, dose==2.0)
```
## Step 3: Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 

### Compare the mean and variance with different methods and applications
```{r compare}
# mean
sapply(split(ToothGrowth$len, ToothGrowth$supp), mean)
sapply(split(dose05$len, dose05$supp), mean)
sapply(split(dose10$len, dose10$supp), mean)
sapply(split(dose20$len, dose20$supp), mean)
# variance
sapply(split(ToothGrowth$len, ToothGrowth$supp), var)
sapply(split(dose05$len, dose05$supp), var)
sapply(split(dose10$len, dose10$supp), var)
sapply(split(dose20$len, dose20$supp), var)

p <- ggplot(aes(x = supp, y = len), data = ToothGrowth) 
p <- p + geom_boxplot(aes(fill = factor(supp))) 
P <- p + xlab("Supplement") 
p <- p + ylab("Tooth Length [mm]") 
P <- p + ggtitle("Tooth Length by Supplement") 
p <- p + geom_point()
p


p2 <- ggplot(ToothGrowth, aes(x=supp, y=len)) 
p2 <- p2 + geom_boxplot(aes(fill=supp))
p2 <- p2 + geom_point()
p2 <- p2 + facet_grid(.~dose)
p2 <- p2 + ggtitle("Tooth Growth by Dosage & Supplement")
p2 <- p2 + xlab("Dosage")
p2 <- p2 + ylab("Tooth Length [mm]")
p2

```

### Detect the Inference by t - test 
#### Test 1: test the difference of supplement
Hypothesis:

$H_0$: same mean of both supplements

$H_a$: different means of both supplements
```{r test1}
t.test(len~supp,paired=FALSE, var.equal=FALSE, data=ToothGrowth)
```
#### Test 2: test the difference under Dosage level 0.5

Hypothesis:

$H_0$: same mean of both supplements

$H_a$: different means of both supplements
```{r test2}
t.test(len~supp,paired=FALSE, var.equal=FALSE, data=dose05)
```
#### Test 3: test the difference under Dosage level 1.0

Hypothesis:

$H_0$: same mean of both supplements

$H_a$: different means of both supplements
```{r test3}
t.test(len~supp,paired=FALSE, var.equal=FALSE, data=dose10)
```
#### Test 4: test the difference under Dosage level 2.0

Hypothesis:

$H_0$: same mean of both supplements

$H_a$: different means of both supplements
```{r test4}
t.test(len~supp,paired=FALSE, var.equal=FALSE, data=dose20)
```
## Step 4: Conclusion
### Supplement Levels overall
As for the supplement OJ and VS, the P-value is 0.06063.

If we set the significant level as 0.05, we would reject the null hypothesis that both supplements have the same means. However, if we make our test not so strict, for example, we set the significant level as 0.1. We would not reject the null hypothesis that both supplements share same mean.

### Supplement Levels separately
We separate the 3 dosage levels, and use t-test to find the difference between supplement levels under the conditions of 3 dosgae levels.

#### Under dosage level of 0.5
The P-value is pretty low, 0.006, which is much lower than 0.05. So, we can think that the OJ and VS supplements are different under dosage level of 0.5.

#### Under dosage level of 1.0
The P-value is pretty low, 0.001, which is much lower than 0.05. So, we can think that the OJ and VS supplements are different under dosage level of 1.0.

#### Under dosage level of 2.0
The P-value is pretty high, 0.95, which is much higher than 0.05. So, we can think that the OJ and VS supplements are not different under dosage level of 2.0.