---
title: 'Part 1: Simulation Exercise Instructions'
Author: Barry Li
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is *1/lambda* and the standard deviation is also *1/lambda*. Set *lambda = 0.2* for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should

1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.

In point 3, focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials.

As a motivating example, compare the distribution of 1000 random uniforms
```{r runif}
# hist(runif(1000))
```
and the distribution of 1000 averages of 40 random uniforms
```{r runif2}
# mns = NULL
# for (i in 1 : 1000) mns = c(mns, mean(runif(40)))
# hist(mns)
```
## Preparation of Simulation
The following code is to setup the parameter of the simulation and distribution. 

You can also embed plots, for example:

```{r parameter}
lambda <- 0.2   # parameter - lambda
n <- 40   # number of exponentials 
sims <- 1000   # number of simulations
rexp(n, lambda)
sim_exp <- replicate(sims, rexp(n, lambda))  #40*1000 matrix

means_exp <- apply(sim_exp, 2, mean) #  mean of 40 numbers in 1000 times
range(means_exp)
hist(means_exp, breaks=40, xlim = c(2,8), main="Means of Exponential Function Simulation", col = "orange")

```

## Step 1: Sample Mean vs. Theoretical Mean 
Theoretically, the mean of exponential distribution is 1/lambda. The theoretical mean should be 1/lambda = 5 .

In the following plot, the red line is the theoretical mean, whereas the blue one is the sample mean.

```{r means}
mean_s <- mean(means_exp) # sample mean
print(paste("mean of sample:", mean_s))
mean_t <- 1/lambda # theoretical mean
print(paste("theoretical mean:", mean_t))

hist(means_exp, main = "Sample Mean vs. Theoretical Mean", xlim =  c(2,8), breaks = 40, xlab = "Simulation Means")
abline(v = mean_s, lwd = "1", col = "blue")
abline(v = mean_t, lwd = "1", col = "red")

```

## Step 2: Sample Variance vs. Theoretical Variance
Theoretically, the variance of exponential distribution is  (1/lambda) / sqrt(n).
```{r SD}
sd_s <- sd(means_exp) # sample standard deviation
print(paste("standard deviation of sample:", sd_s^2))
sd_t <- (1/lambda)/sqrt(n) # theoretical standard deviation
print(paste("theoretical standard deviation:", sd_t^2))

```
## step 3: Distribution
In this step, the exponential distribution would be compared with CLT(central limit theorem):the means of the sample simulations should follow a normal distribution.
The blue line is normal distribution.
The red line is the 
```{r distribution}
#General Plot with ditribution curve drawn
hist(means_exp, prob=TRUE, col="aliceblue", main="Distribution Comparation", breaks=40, xlim=c(2,8), xlab = "Simulation Means")
lines(density(means_exp), lwd=3, col="red")

# Normal distribution line creation
x <- seq(min(means_exp), max(means_exp), length=n)
y <- dnorm(x, mean=1/lambda, sd=(1/lambda)/sqrt(n))
lines(x, y, pch=22, col="blue", lwd=3)
```